Fourier-domain modulations and delays of gravitational-wave signals

### 2018-03-01 docs/paper_method.tex
Changes to the noise and utility executables.
- updated noise to include the 2017 proposal noise used in the LDC (NOTE: made the new default)
- updated interface for ComputeLISASNR and GenerateTDIFD
More edits through IV.B

\jgb{In the case of precessing binaries, applying the SPA directly to the modulated signal is also prone to pathologies. In Ref.~\cite{KCY13,KCY14}, a formalism (called shifted uniform asymptotics or SUA) was introduced to go beyond the SPA and compute more accurately the modulation in the Fourier domain; however, this formalism still relies on the SPA for the underlying precessing-frame signal, and is as such limited to inspiraling signals. \textbf{[JB: Still uncertain about the rest of this paragraph.]} Most directly, a time-to-frequency correspondence, such as the one given by the SPA, can be exploited to directly carry over the precision modulation to the Fourier domain.} This corresponds, in a sense to be detailed below, to the zeroth-order approximation of the SUA, and it is used implicitly\footnote{With the difference that the expressions used for the modulation are given in terms of frequency directly for the inspiral, and are then extended to also cover high frequencies in the merger-ringdown part of the signal, as we will discuss below.} for the simplified precession of the phenomenological waveforms of Ref.~\cite{Hannam+13}.

@@

We begin by detailing the model that we use for the response of a LISA-like detector, together with the assumptions used and their limitations. Since our aim is to assess the accuracy of our direct Fourier-domain treatment of the response, we can focus on the gravitational-wave contribution to the basic observables. We therefore use a somewhat simplified model for the time-domain response, ignoring corrections that are crucial from the point of view of noise cancellations, keeping in mind that the response model can be enriched later without affecting the conclusions of the present analysis.

@@

Additionally, we limit ourselves to a rigid model for the orbits of the constellation, namely we assume that the constellation remains in an equilateral configuration with fixed armlengths $L=2.5\times 10^{6}\mathrm{km}$ (the latest configuration~\cite{LISA17}). This simplified orbits neglect (iv) the effect of excentricity of the individual Keplerian orbits, which introduces corrections at the level $e^{2}$ (v) the effect of gravitational perturbations coming from other celestial bodies, such as the Earth, the quadrupole of the Sun, and the other planets.

@@

It is natural to split the response~\eqref{eq:yslr} into two steps: the orbital delay related to the orbit around the Sun of the whole constellation on one side, and the constellation response on the other side. The baselines for the delays are indeed very different in the two cases. Geometrical projection factors aside, we have for the orbit around the Sun $R=1\text{au}=1.5\times 10^{8} \text{km}$, while the detector armlength is $L=2.5\times 10^{6}\text{km}$ (in the configuration proposed in~\cite{LISA17}). It is natural to define two transfer frequencies for the two relevant length scales for the delays, defined such that a wavelength fits within this length scale, i.e. $2\pi f d = 1$. This gives
\begin{subequations}\label{eq:transferfrequencies}
\begin{align}
	f_{R} &= 3.2\times10^{-4}\Hz \,,\\
	f_{L} &= 1.9\times 10^{-2}\Hz \,.
\end{align}
\end{subequations}
The LISA response behaves qualitatively differently on the frequency bands separated by these transfer frequencies.

The first step consists simply in applying the varying time delay to bring the signal from the SSB to the center of the LISA triangular constellation, common to all $y_{slr}$ observables. For $h^{\rm TT}$ the transverse-traceless gravitational waveform in matrix form, we write this orbital time delay as
\be\label{eq:defresponse0}
	h_{0}^{\rm TT} (t) = h^{\rm TT}(t-\hatk\cdot p_{0}) \,,
\ee
with $p_{0}$ the position of the constellation center, which follows the Earth orbit around the Sun. The second step consists in the remaining, constellation-centered response, which we write as
\begin{align}\label{eq:defresponseL}
	y_{slr} &= \frac{1}{2} \frac{1}{1 - \hatk\cdot n_{l}} \nn\\
	& \cdot n_{l}\cdot \left[ h_{0}^{\rm TT}(t - L - \hatk\cdot p^{L}_{s}) - h_{0}^{\rm TT}(t - \hatk\cdot p^{L}_{r}) \right] \cdot n_{l}\,,
\end{align}
where we introduced the positions of the spacecraft relative to the center of the constellation, $p^{L}_{A} \equiv p_{A} - p_{0}$.

@@

We now turn to the transposition of~\eqref{eq:defresponse0} and~\eqref{eq:defresponseL} to the Fourier domain. Applying a pure delay as in~\eqref{eq:defresponse0} translates into
\be\label{eq:G0}
	G_{0}(f, t) = e^{-2i\pi f d_{0}(t)} \,,
\ee
with $d_{0} = -\hatk \cdot p_{0}$ the delay associated to the orbit around the Sun. For the leading-order response~\eqref{eq:transferlocal}, this gives a Fourier-domain transfer function common to all modes, that is a pure phase factor, proportional to the frequency but also $t_{f}$-dependent:
\be\label{eq:transfer0local}
	\calT_{0}^{\rm local}(f) = G_{0}(f, \tf)\,.
\ee
If $(\lambda, \beta)$ are the ecliptic longitude and latitude of the source in the sky, and if the orbital phase is set by convention to $0$ at $t=0$, the orbital delay has the simple expression
\be\label{eq:delay0}
	d_{0}(t) = -R \cos\beta \cos\left(\Omega_{0}t - \lambda\right)\,.
\ee

For the constellation response~\eqref{eq:defresponseL}, we can separate the two delays by writing
\begin{align}\label{eq:decomposeGslr}
	F_{slr}^{L}(t) &= \frac{1}{2} \frac{1}{1 - \hatk\cdot n_{l}(t)} n_{l}(t) \cdot P_{\ell m} \cdot n_{l} (t) \,,\nn\\
	d_{s,L}(t) &= -L - k\cdot p_{s}^{L}(t) \,, \quad d_{r}(t) = - k\cdot p_{r}^{L}(t) \,,\nn\\
	G_{slr}^{L}(f,t) &=  F_{slr}^{L}(t) \left( e^{-2i\pi f d_{s,L}(t)} - e^{-2i\pi f d_{r}(t)} \right) \,.
\end{align}
When also using the rigid approximation for the constellation, where the armlengths are fixed, a particular simplification occurs when combining these individual delays, thanks to the relation $p^{L}_{r} - p^{L}_{s} =  L n_{l}$:
\begin{align}\label{eq:GslrL}
	G_{slr}^{L}(f,t) &= \frac{i \pi f L}{2} \sinc \left[ \pi f L\left(1-\hatk\cdot n_{l} \right) \right] \nn\\
	& \quad \cdot \exp\left[ i \pi f \left( L + \hatk\cdot \left( p_{1}^{L} + p_{2}^{L} \right) \right) \right]  n_{l} \cdot P_{\ell m} \cdot n_{l} \,,
\end{align}
with all time-dependent vectors evaluated at $t$. In the local approximation~\eqref{eq:transferlocal}, the Fourier-domain transfer function then reads
\begin{align}\label{eq:transferLlocal}
	\calT_{slr}^{L, \mathrm{local}}(f) &= G_{slr}^{L}(f, \tf) \,.
\end{align}
This leading-order result~\eqref{eq:transfer0local}-\eqref{eq:transferLlocal} is familiar, as it was derived in the past taking the point of view of monochromatic signals (see e.g.~\cite{} \SM{[C]}).

Note that, if the corrections of Sec.~\ref{subsec:delays} are included for the constellation delays, the transfer function will not have this simple form anymore, as $\dot{d}$ will have a different velocity-dependent expression for the sending and receiving spacecraft. One must then handle separately $d_{s,L}$ and $d_{r}$ in~\eqref{eq:decomposeGslr}.

The response~\eqref{eq:transfer0local}-\eqref{eq:delay0} takes a simple analytic form, but the phase contribution of this delay is significant on most of the frequency band and can be large for $f \gg f_{R}$.

The response~\eqref{eq:GslrL}-\eqref{eq:transferLlocal} can be interpreted as the Fourier-domain translation of a discrete derivative taken on the waveform. The factor shows that the amplitude of the response is $\propto f$ in the low-frequency limit $f\ll f_{L}$, where the other factors are essentially unity. For $f\gtrsim f_{L}$, the $\sinc$ and the phase of the exponential give raise to additional structure in the response, including zero-crossings when the projected armlength corresponds to a wavelength. From~\eqref{eq:transferLlocal}, an expansion for small $f\ll f_{L}$ yields back a Fourier-domain analog of the low-frequency approximation of the response~\cite{Cutler97, RCP04}, which amounts to having two LIGO-type interferometers turned by $\pi/4$ and set in motion. See Fig.~\ref{} and ~\ref{} \SM{[R]} for examples of LISA transfer functions.

For the gravitational waveform itself, we will use the PhenomD model~\cite{Khan+15,Husa+15}, which provides Fourier-domain inspiral-merger-ringdown waveforms for aligned spins.

@@

\begin{figure}
  \centering
  \includegraphics[width=.9\linewidth]{plots/lisafom_Psi_McDeltat.pdf}
  \caption{\SM{[To be checked and redone]} Contour levels for the analytical estimate of the error measure $\epsilon_{\Psi}$ at the starting frequency, as a function of chirp mass $\Mchirp$ and time to coalescence $\Delta t$. Blue corresponds to the orbital response and red to the constellation response. The colored shaded areas indicate regions where $\epsilon_{\Psi} \geq 1$, where the perturbative formalism is expected to break down. In the gray area, $f_{\rm start}$ given in~\eqref{eq:fstartN} is lower than the lowest in-band frequency $f_{\rm min} = 10^{-5}\Hz$, so that the signal starts at $f_{\rm min}$ and $\epsilon_{\Psi}$ becomes independent of $\Delta t$.}
  \label{fig:lisafomPsiMcDeltat}
\end{figure}

@@

We will now use the approximate error measures $\epsilon$ introduced in~\eqref{eq:deffom} to estimate, for each type of correction (phase, amplitude, delay), the level of errors in the transfer function induced by ignoring this correction.

To obtain an order-of-magnitude estimate for the error measures $\epsilon$~\eqref{eq:deffom}, we will use the Newtonian-order expressions~\eqref{eq:timescalesN} for the signal-dependent timescales $\Tf$ and $T_{A1}$. Higher-order amplitude terms beyond the first one in~\eqref{eq:resulttaylorA}, as well as phase terms beyond the second derivative in~\eqref{eq:expandPsi}, will turn out to be completely negligible in the LISA case and we will ignore them in the following. It is useful to separate the orbital response~\eqref{eq:transfer0local} and the constellation response~\eqref{eq:transferLlocal}, as the different baseline of the delays (orbital radius $R$ or armlength $L$) as well as the presence of a time-varying prefactor $F(t)$ both affect the result. When estimating the magnitude of the relevant derivatives of $G$, we have to take into account dimensionless delay factors of the type $2\pi f d$.

We start with the orbital response, which takes the form of a pure delay $G_{0}(f, t) = e^{-2 i \pi f d_{0}(t)}$, and obtain
\begin{subequations}
\begin{align}
	\frac{1}{G_{0}} \partial_{t} G_{0} &= -2i\pi f \dot{d}_{0}\,,\\
	\frac{1}{G_{0}} \partial_{tt} G_{0} &= -2i\pi f \ddot{d}_{0} - 4\pi^{2} f^{2} \dot{d}_{0}^{2} \,,\\
	\frac{1}{G_{0}} \partial_{tf} G_{0} &= -2 i \pi \dot{d}_{0} - 4\pi^{2} f d_{0} \dot{d}_{0} \,.
\end{align}
\end{subequations}

Since the one-arm constellation response~\eqref{eq:decomposeGslr}-\eqref{eq:transferLlocal} is analogous to a discrete time derivative of the signal, it is appropriate to keep explicit an overall factor $f$ reflecting this structure. Hence we write symbolically $G_{L}(f,t) \sim f F(t) e^{-2i \pi f d_{L}(t)}$, where $d_{L}$ represents a delay term and $F(t)$ represents the rest of the geometric factors in~\eqref{eq:transferLlocal} where we ignore the $f$-dependence. This gives
\begin{subequations}
\begin{align}
	\frac{1}{G_{L}} \partial_{t} G_{L} &\sim -2i\pi f \dot{d}_{c} + \frac{\dot{F}}{F}\,,\\
	\frac{1}{G_{L}} \partial_{tt} G_{L} &\sim -2i\pi f \ddot{d}_{c} - 4\pi^{2} f^{2} \dot{d}_{c}^{2} - 2i\pi f \dot{d}_{c} \frac{\dot{F}}{F} - \frac{\dot{F}^{2}}{F^{2}} + \frac{\ddot{F}}{F} \,,\\
	\frac{1}{G_{L}} \partial_{tf} G_{L} &\sim -4 i \pi \dot{d}_{c} - 2i\pi d_{c} \frac{\dot{F}}{F} - 4\pi^{2} f d_{c} \dot{d}_{c} + \frac{1}{f}\frac{\dot{F}}{F}\,.
\end{align}
\end{subequations}

@@

An important point in the above is the presence of delay factors in the forms of powers of $2\pi f d$. Thus, the magnitude of the corrections cannot be evaluated with the simple replacement $\partial_{t}^{n} \rightarrow \Omega_{0}^{n}$, and the frequency dependency of the error measures $\epsilon$ will depend on the frequency being above or below the transfer frequencies $f_{R}$ and $f_{L}$ defined in~\eqref{eq:transferfrequencies}. The error measure $\epsilon_{d}$ does not depend on signal-dependent timescales and can be directly read off the estimates above, with $\epsilon_{d} = |\partial_{tf}G/(2\pi G)|$. For the error measures $\epsilon_{\Psi}$ and $\epsilon_{A1}$, we can combine the above derivatives with the Newtonian timescales given in~\eqref{eq:timescalesN} for the inspiral phase of the signal.

The starting frequency of the signal will play an important role. As a function of the time remaining before merger $\Delta t$, from the Newtonian relation~\eqref{eq:omegaphiN} we have (using the chirp mass $\Mchirp = M \nu^{3/5}$):
\be\label{eq:fstartN}
	f_{\rm start} = 1.75\times 10^{-5} \Hz \left( \frac{\Mchirp}{10^{6}M_{\odot}} \right)^{-5/8} \left( \frac{\Delta t}{10 \yr} \right)^{-3/8} \,.
\ee
This gives in turn for the Newtonian timescales~\eqref{eq:timescalesN}
\begin{subequations}\label{eq:timescalesNfstart}
\begin{align}
	\Tf^{\rm N} &= 8.78\times10^{-2}\yr \left( \frac{\Mchirp}{10^{6}M_{\odot}} \right)^{5/16} \left( \frac{\Delta t}{10 \yr} \right)^{11/16} \,, \\
	T_{A1}^{\rm N} &= 3.37\times10^{-4}\yr \left( \frac{\Mchirp}{10^{6}M_{\odot}} \right)^{5/8} \left( \frac{\Delta t}{10 \yr} \right)^{3/8} \,.
\end{align}
\end{subequations}
Note that if we consider that the frequency band of the instrument starts at a minimal frequency $f_{\rm min}$, for sufficiently high $\Mchirp$ this entry in the frequency band will mark the beginning of the signal, instead of $f_{\rm start}$ which becomes irrelevant. For $f_{\rm min} = 10^{-5}\Hz$ and $\Delta t \leq 10\yr$, this is the case for $\Mchirp \geq 2.45\times 10^{6}\Msol $. Thus, our final analytical estimates for the $\epsilon$ error measures are built by inserting~\eqref{eq:timescalesNfstart} and~\eqref{eq:estimatederivorb}-\eqref{eq:estimatederivconst} in~\eqref{eq:deffom}, while ensuring the frequency cut $10^{-5}\Hz \leq f \leq 1\Hz$.

An important distinction needs to be made between what we will call merging binaries, systems close to coalescence that will merge during the LISA mission lifetime or a few years later, and slowly-chirping binaries still in the deep inspiral phase, of which we observe only a small snapshot in Fourier domain as they do not sweep to the end of the frequency band. The massive black hole binaries (MBH) that will be observed by LISA~\cite{LISA17} fall within the first category, with their merger in band, while the proposed population of stellar-origin black hole binaries (SOBH)~\cite{Sesana16}, with masses comparable to the LIGO/Virgo detections, will comprise both merging binaries, i.e. exiting the LISA band towards larger frequencies during observations, and slowly-chirping binaries hundereds or thousands of years away from merger. Galactic binaries (GB) fall entirely within the second category, and are very close to monochromatic over the lifetime of LISA~\cite{LISA17} \SM{[C: complete]}.

We turn first to the slowly-chirping binaries. For these systems, we will focus on $\epsilon_{\Psi}$, which will be the most important error measure. Fig.~\ref{fig:lisafomPsiMcDeltat} shows contour levels for the value of our simple analytical estimate for $\epsilon_{\Psi}$ at the beginning of the observations, as a function of the chirp mass and time to merger, for both the orbital and the constellation response. The limit $\epsilon_{\Psi} = 1$ is used to single out areas where the perturbative treatment of Sec.~\ref{sec:formalism} is expected to break down, although the precise location of this boundary will vary, depending on the accuracy level required and on the orientation angles. We also single out the high-mass region where the start of the signal is set by its entry in the frequency band, taken to be $10^{-5}\Hz$, so that there is no dependence on the time to merger anymore. We find that, in the LIGO/Virgo mass range $10^{1}-10^{2} \Msol$, we reach the limit $\epsilon_{\Psi}=1$ first for the orbital response, for time-to-merger values within the expected observed range~\cite{Sesana16}. We will introduce in Sec~\ref{subsec:comblisa} an alternative approach to deal with those signals. For completeness, we also show in Fig.~\ref{fig:lisafomPsiMcDeltat} the result for intermediate and massive black hole systems, although we expect to observe mainly merging systems for this mass range.

>>>>

\jgb{In the case of precessing binaries, applying the SPA directly to the modulated signal is also prone to pathologies. In Ref.~\cite{KCY13,KCY14}, a formalism (called shifted uniform asymptotics or SUA) was introduced to go beyond the SPA and compute more accurately the modulation in the Fourier domain; however, this formalism still relies on the SPA for the underlying precessing-frame signal, and is as such limited to inspiraling signals. \textbf{[JB: Did I capture your thoughts correctly in the following.]}
The simplified treatment of the precession response in the phenomenological waveforms in Ref.~\cite{Hannam+13} takes another approach, treating the precession modulation in the frequency domain by directly associating the Fourier frequency with (a multiple of) the post-Newtonain orbital frequency. As we will explain, this corresponds to the zeroth-order approximation of the SUA.}

@@

We begin by detailing the model that we use for the response of a LISA-like detector, together with the assumptions used and their limitations. Since our aim is to assess the accuracy of our direct Fourier-domain treatment of the response, we can focus on the gravitational-wave contribution to the basic single-linkobservables. We therefore use a somewhat simplified model for the time-domain response, ignoring corrections that are crucial from the point of view of noise cancellations, keeping in mind that the response model can be enriched later without affecting the conclusions of the present analysis.

@@

Additionally, we limit ourselves to a rigid model for the orbits of the constellation, namely we assume that the constellation remains in an equilateral configuration with fixed armlengths. These simplified orbits neglect (iv) effects of order $e^{2}$ from the eccentricity of the individual Keplerian orbits, (v) the effect of gravitational perturbations coming from other celestial bodies, such as the Earth, the quadrupole of the Sun, and the other planets.

@@

It is natural to split the response~\eqref{eq:yslr} into two steps: first the orbital delay related to the orbit around the Sun of the whole constellation, and then the constellation response. The baselines for the delays are indeed very different in the two cases. Geometrical projection factors aside, we have for the orbit around the Sun $R=1\text{au}=1.5\times 10^{8} \text{km}$, while the detector armlength is $L=2.5\times 10^{6}\text{km}$ (in the configuration proposed in~\cite{LISA17}). It is natural to define two transfer frequencies for the two relevant length scales for the delays, defined such that a wavelength fits within this length scale, i.e. $2\pi f d = 1$. This gives
\begin{subequations}\label{eq:transferfrequencies}
\begin{align}
	f_{R} &= 3.2\times10^{-4}\Hz \,,\\
	f_{L} &= 1.9\times 10^{-2}\Hz \,.
\end{align}
\end{subequations}
The LISA response behaves qualitatively differently on the frequency bands separated by these transfer frequencies. \jgb{[\bf I don't understand the meaning or intent of the previous sentence]}

\jgb{The first stage of the response, the orbital delay, consists simply in applying the varying time delay to bring the wavefront sampling point from the SSB reference to the center} of the LISA triangular constellation, common to all $y_{slr}$ observables. For $h^{\rm TT}$ the transverse-traceless gravitational waveform in matrix form, we write this orbital time delay as
\be\label{eq:defresponse0}
	h_{0}^{\rm TT} (t) = h^{\rm TT}(t-\hatk\cdot p_{0}) \,,
\ee
with $p_{0}$ the position of the constellation center, which follows the Earth orbit around the Sun. The second \jgb{stage of the response calculation comprises the remaining, constellation-centered response. For the single-link contribution to the response, for the laser link from spacecraft $s$ to spacecraft $r$ along a path in direction $n_l=$ [{\bf Check that I got the notation right}], we write}
\begin{align}\label{eq:defresponseL}
	y_{slr} &= \frac{1}{2} \frac{1}{1 - \hatk\cdot n_{l}} \nn\\
	& \cdot n_{l}\cdot \left[ h_{0}^{\rm TT}(t - L - \hatk\cdot p^{L}_{s}) - h_{0}^{\rm TT}(t - \hatk\cdot p^{L}_{r}) \right] \cdot n_{l}\,,
\end{align}
where we reference the positions of the spacecraft relative to the center of the constellation, $p^{L}_{A} \equiv p_{A} - p_{0}$.

@@

We now turn to the transformation of~\eqref{eq:defresponse0} and~\eqref{eq:defresponseL} to the Fourier domain. Applying a pure delay as in~\eqref{eq:defresponse0} translates into
\be\label{eq:G0}
	G_{0}(f, t) = e^{-2i\pi f d_{0}(t)} \,,
\ee
with $d_{0} = -\hatk \cdot p_{0}$ the delay associated to the orbit around the Sun. For the leading-order response~\eqref{eq:transferlocal}, this gives a Fourier-domain transfer function common to all modes, that is a pure phase factor, proportional to the frequency but also $t_{f}$-dependent:
\be\label{eq:transfer0local}
	\calT_{0}^{\rm local}(f) = G_{0}(f, \tf)\,.
\ee
If $(\lambda, \beta)$ are the ecliptic longitude and latitude of the source in the sky, and if the orbital phase is set by convention to $0$ at $t=0$, the orbital delay has the simple expression
\be\label{eq:delay0}
	d_{0}(t) = -R \cos\beta \cos\left(\Omega_{0}t - \lambda\right)\,.
\ee

For the constellation response~\eqref{eq:defresponseL}, we can \jgb{separately treat the orbital delay $d_r$ and the constellation delay at spacecraft $s$, $d_{s,L}$}, by writing
\begin{align}\label{eq:decomposeGslr}
	F_{slr}^{L}(t) &= \frac{1}{2} \frac{1}{1 - \hatk\cdot n_{l}(t)} n_{l}(t) \cdot P_{\ell m} \cdot n_{l} (t) \,,\nn\\
	d_{s,L}(t) &= -L - k\cdot p_{s}^{L}(t) \,, \quad d_{r}(t) = - k\cdot p_{r}^{L}(t) \,,\nn\\
	G_{slr}^{L}(f,t) &=  F_{slr}^{L}(t) \left( e^{-2i\pi f d_{s,L}(t)} - e^{-2i\pi f d_{r}(t)} \right) \,.
\end{align}
\jgb{Since we also assume the {\bf[right]} } rigid approximation for the constellation, where the armlengths are fixed, a particular simplification occurs when combining these individual delays, thanks to the relation $p^{L}_{r} - p^{L}_{s} =  L n_{l}$:
\begin{align}\label{eq:GslrL}
	G_{slr}^{L}(f,t) &= \frac{i \pi f L}{2} \sinc \left[ \pi f L\left(1-\hatk\cdot n_{l} \right) \right] \nn\\
	& \quad \cdot \exp\left[ i \pi f \left( L + \hatk\cdot \left( p_{1}^{L} + p_{2}^{L} \right) \right) \right]  n_{l} \cdot P_{\ell m} \cdot n_{l} \,,
\end{align}
with all time-dependent vectors evaluated at $t$. In the local approximation~\eqref{eq:transferlocal}, the Fourier-domain transfer function then reads
\begin{align}\label{eq:transferLlocal}
	\calT_{slr}^{L, \mathrm{local}}(f) &= G_{slr}^{L}(f, \tf) \,.
\end{align}
This leading-order result~\eqref{eq:transfer0local}-\eqref{eq:transferLlocal} is familiar, as it was derived in the past taking the point of view of monochromatic signals (see e.g.~\cite{} \SM{[C]}).

Note that, if the corrections of Sec.~\ref{subsec:delays}, \jgb{for non-negligible $\dot d$}, are included for the constellation delays, the transfer function will not have this simple form anymore, as $\dot{d}$ will have a different velocity-dependent expression for the sending and receiving spacecraft. One must \jgb{separately handle $d_{s,L}$ and $d_{r}$ in~\eqref{eq:decomposeGslr} to compute the corrections.}

The \jgb{orbital} response~\eqref{eq:transfer0local}-\eqref{eq:delay0} takes a simple analytic form, but the phase contribution of this delay is significant across most of the frequency band and can be large for $f \gg f_{R}$.

The \jgb{constellation} response~\eqref{eq:GslrL}-\eqref{eq:transferLlocal} can be interpreted as the Fourier-domain translation of a discrete derivative taken on the waveform. The \jgb{leading factor in~\eqref{eq:GslrL}} shows that the amplitude of the response is proportional to $f$ in the low-frequency limit $f\ll f_{L}$, where the other factors are essentially unity. For $f\gtrsim f_{L}$, the $\sinc$ and the phase of the exponential generate additional structure in the response, including zero-crossings when the projected armlength \jgb{is an integer number of wavelengths}. From~\eqref{eq:transferLlocal}, an expansion for small $f\ll f_{L}$ yields back a Fourier-domain analog of the low-frequency approximation of the response~\cite{Cutler97, RCP04}, which \jgb{is equivalent to} having two LIGO-type interferometers turned by $\pi/4$ and set in motion. See Fig.~\ref{} and ~\ref{} \SM{[R]} for examples of LISA transfer functions.

\jgb{For analysis of the response we need a concrete set of gravitational waveforms. We will} use the PhenomD model~\cite{Khan+15,Husa+15}, which provides Fourier-domain inspiral-merger-ringdown waveforms for aligned spins.

@@

\begin{figure}
  \centering
  \includegraphics[width=.9\linewidth]{plots/lisafom_Psi_McDeltat.pdf}
  \caption{\SM{[To be checked and redone]} Contour levels for the analytical estimate of the error measure $\epsilon_{\Psi}$ at the starting frequency, as a function of chirp mass $\Mchirp$ and time to coalescence $\Delta t$. Blue corresponds to the orbital response and red to the constellation response. The colored shaded areas indicate regions where $\epsilon_{\Psi} \geq 1$, where the perturbative formalism is expected to break down. In the gray area, $f_{\rm start}$ given in~\eqref{eq:fstartN} is lower than the lowest in-band frequency $f_{\rm min} = 10^{-5}\Hz$, so that the signal starts at $f_{\rm min}$ and $\epsilon_{\Psi}$ becomes independent of $\Delta t$.\jgb{\bf[Maybe not shade the ``grey area'' since it looks excluded, instead can call it ``the region to the right of the black line.'']} }
  \label{fig:lisafomPsiMcDeltat}
\end{figure}

@@

Using the approximate error measures $\epsilon$ introduced in~\eqref{eq:deffom}, we will now estimate, for each type of correction (phase, amplitude, delay), the size of errors in the transfer function.

To obtain an order-of-magnitude estimate for the error measures $\epsilon$~\eqref{eq:deffom}, we will use the Newtonian-order expressions~\eqref{eq:timescalesN} for the signal-dependent timescales $\Tf$ and $T_{A1}$. Higher-order amplitude terms beyond the first one in~\eqref{eq:resulttaylorA}, as well as phase terms beyond the second derivative in~\eqref{eq:expandPsi}, will turn out to be completely negligible in the LISA case and we will ignore them in the following. It is useful to separate the orbital response~\eqref{eq:transfer0local} and the constellation response~\eqref{eq:transferLlocal}, as the different baseline of the delays (orbital radius $R$ or armlength $L$) as well as the presence of a time-varying prefactor $F(t)$ both affect the result. When estimating the magnitude of the relevant derivatives of $G$, we must also take into account dimensionless delay factors of the type $2\pi f d$.

We start with the orbital response, which takes the form of a pure delay $G_{0}(f, t) = e^{-2 i \pi f d_{0}(t)}$ \jgb{\bf[Why subscripted with zero not O or R?]}, and obtain
\begin{subequations}
\begin{align}
	\frac{1}{G_{0}} \partial_{t} G_{0} &= -2i\pi f \dot{d}_{0}\,,\\
	\frac{1}{G_{0}} \partial_{tt} G_{0} &= -2i\pi f \ddot{d}_{0} - 4\pi^{2} f^{2} \dot{d}_{0}^{2} \,,\\
	\frac{1}{G_{0}} \partial_{tf} G_{0} &= -2 i \pi \dot{d}_{0} - 4\pi^{2} f d_{0} \dot{d}_{0} \,.
\end{align}
\end{subequations}

Since the one-arm constellation response~\eqref{eq:decomposeGslr}-\eqref{eq:transferLlocal} is analogous to a discrete time derivative of the signal, it is appropriate to keep explicit an overall factor $f$ reflecting this structure. Hence we write symbolically $G_{L}(f,t) \sim f F(t) e^{-2i \pi f d_{L}(t)}$, where $d_{L}$ represents a delay term and $F(t)$ represents the rest of the geometric factors in~\eqref{eq:transferLlocal}, \jgb{for which we momentarily} ignore the $f$-dependence. This gives
\begin{subequations}
\begin{align}
	\frac{1}{G_{L}} \partial_{t} G_{L} &\sim -2i\pi f \dot{d}_{c} + \frac{\dot{F}}{F}\,,\\
	\frac{1}{G_{L}} \partial_{tt} G_{L} &\sim -2i\pi f \ddot{d}_{c} - 4\pi^{2} f^{2} \dot{d}_{c}^{2} - 2i\pi f \dot{d}_{c} \frac{\dot{F}}{F} - \frac{\dot{F}^{2}}{F^{2}} + \frac{\ddot{F}}{F} \,,\\
	\frac{1}{G_{L}} \partial_{tf} G_{L} &\sim -4 i \pi \dot{d}_{c} - 2i\pi d_{c} \frac{\dot{F}}{F} - 4\pi^{2} f d_{c} \dot{d}_{c} + \frac{1}{f}\frac{\dot{F}}{F}\,.
\end{align}
\end{subequations}

@@

\jgb{\bf[This paragraph is confusing.  Can we point concretely to these factors somewhere, (e.g. in the expansion of the exponential.xxx)? Also it seems to say that we can't do what we just did. Should discuss.]} An important point in the above is the presence of delay factors in the forms of powers of $2\pi f d$. Thus, the magnitude of the corrections cannot be evaluated with the simple replacement $\partial_{t}^{n} \rightarrow \Omega_{0}^{n}$, and the frequency dependency of the error measures $\epsilon$ will depend on the frequency being above or below the transfer frequencies $f_{R}$ and $f_{L}$ defined in~\eqref{eq:transferfrequencies}. The error measure $\epsilon_{d}$ does not depend on signal-dependent timescales and can be directly read off the estimates above, with $\epsilon_{d} = |\partial_{tf}G/(2\pi G)|$. For the error measures $\epsilon_{\Psi}$ and $\epsilon_{A1}$, we can combine the above derivatives with the Newtonian timescales given in~\eqref{eq:timescalesN} for the inspiral phase of the signal.

The starting frequency of the signal will play an important role. As a function of the time remaining before merger $\Delta t$, from the Newtonian relation~\eqref{eq:omegaphiN} we have (using the chirp mass $\Mchirp = M \nu^{3/5}$):
\be\label{eq:fstartN}
	f_{\rm start} = 1.75\times 10^{-5} \Hz \left( \frac{\Mchirp}{10^{6}M_{\odot}} \right)^{-5/8} \left( \frac{\Delta t}{10 \yr} \right)^{-3/8} \,.
\ee
This gives in turn for the Newtonian timescales~\eqref{eq:timescalesN}
\begin{subequations}\label{eq:timescalesNfstart}
\begin{align}
	\Tf^{\rm N} &= 8.78\times10^{-2}\yr \left( \frac{\Mchirp}{10^{6}M_{\odot}} \right)^{5/16} \left( \frac{\Delta t}{10 \yr} \right)^{11/16} \,, \\
	T_{A1}^{\rm N} &= 3.37\times10^{-4}\yr \left( \frac{\Mchirp}{10^{6}M_{\odot}} \right)^{5/8} \left( \frac{\Delta t}{10 \yr} \right)^{3/8} \,.
\end{align}
\end{subequations}
Note that if we \jgb{think of LISA as effectively insensitive below some minimal frequency $f_{\rm min}$, then for sufficiently high $\Mchirp$ this point of entry in the sensitive band will mark the beginning of the signal, obviating the relevance of $f_{\rm start}$}. For $f_{\rm min} = 10^{-5}\Hz$ and $\Delta t \leq 10\yr$, this is the case for $\Mchirp \geq 2.45\times 10^{6}\Msol $. Thus, our final analytical estimates for the $\epsilon$ error measures are built by inserting~\eqref{eq:timescalesNfstart} and~\eqref{eq:estimatederivorb}-\eqref{eq:estimatederivconst} in~\eqref{eq:deffom}, while ensuring the frequency cut $10^{-5}\Hz \leq f \leq 1\Hz$.

\jgb{It is useful to distinguish} between what we will call merging binaries, systems close to coalescence that will merge during the LISA mission lifetime or a few years later, and slowly-chirping binaries still in the deep inspiral phase, of which we observe only a small snapshot in Fourier domain as they do not sweep to the end of the frequency band. The massive black hole binaries (MBH) that will be observed by LISA~\cite{LISA17} fall within the first category, with their merger in band, while the proposed population of stellar-origin black hole binaries (SOBH)~\cite{Sesana16}, with masses comparable to the LIGO/Virgo detections, will comprise both merging binaries, i.e. exiting the LISA band towards larger frequencies during observations, and slowly-chirping binaries hundereds or thousands of years away from merger. Galactic binaries (GB) fall entirely within the second category, and are very close to monochromatic over the lifetime of LISA~\cite{LISA17} \SM{[C: complete]}.

We turn first to the slowly-chirping binaries. For these systems, we will focus on $\epsilon_{\Psi}$, which will be the most important error measure. Fig.~\ref{fig:lisafomPsiMcDeltat} shows contour levels for the value of our simple analytical estimate for $\epsilon_{\Psi}$ at the beginning of the observations, as a function of the chirp mass and time to merger, for both the orbital and the constellation response. The limit $\epsilon_{\Psi} = 1$ is used to single out areas where the perturbative treatment of Sec.~\ref{sec:formalism} is expected to break down, although the precise location of this boundary will vary, depending on the accuracy level required and on the orientation angles. We also single out the high-mass region where the start of the signal is set by its entry \jgb{into the sensitive band, at $f_{\rm min}$ making the error measure independent of time to merger}. We find that, for binaries in the LIGO/Virgo mass range $10^{1}-10^{2} \Msol$, we reach the limit $\epsilon_{\Psi}=1$ first for the orbital response, for time-to-merger values within the expected observed range~\cite{Sesana16}. We will introduce in Sec~\ref{subsec:comblisa} an alternative approach to deal with those signals. For completeness, we also show in Fig.~\ref{fig:lisafomPsiMcDeltat} the result for intermediate and massive black hole systems, although we expect to observe mainly merging systems for this mass range.
